### 2025-08-02 output/tables/lead_by_players.tex
Re-organize file structure (#3)
* update file paths for data CSVs

* add input/output dirs to filepaths in generate_results.R

* rename latex source files

* ignore input/output and latex auxiliary files

* update file paths in latex code

* fix bug with underscore in url

* gitignore *.blg and *.synctex.gz

* move tracked figures and tables to output folder

* update folder structure in README

* move data wrangling code for first figure to generate_results.R

* remove code for unused runners_outs_two_agent table

* update output filepath for lead_by_runners_outs.tex

* fix style in generate_

% latex table generated in R 4.3.1 by xtable 1.8-4 package
% Sun Jul 21 17:31:33 2024
 10th Percentile Battery & 10th Percentile Runner & 10.5 & 11.9 & 14.5 \\ 
  10th Percentile Battery & Median Runner & 11.3 & 12.5 & 14.7 \\ 
  10th Percentile Battery & 90th Percentile Runner & 12.0 & 13.0 & 14.8 \\ 
  Median Battery & 10th Percentile Runner & 9.6 & 11.3 & 14.1 \\ 
  Median Battery & Median Runner & 10.3 & 11.7 & 14.1 \\ 
  Median Battery & 90th Percentile Runner & 11.0 & 12.1 & 14.2 \\ 
  90th Percentile Battery & 10th Percentile Runner & 7.8 & 9.8 & 12.6 \\ 
  90th Percentile Battery & Median Runner & 8.2 & 10.0 & 12.7 \\ 
  90th Percentile Battery & 90th Percentile Runner & 8.8 & 10.4 & 12.8 \\