### 2025-08-02 output/tables/lead_by_players.tex
switch paper to new tables

\begin{tabular}{cc|rrr}
\multicolumn{2}{c|}{Skill Percentile} & \multicolumn{3}{c}{Disengagements}\\
Battery & Runner & 0 & 1 & 2 \\
% latex table generated in R 4.5.0 by xtable 1.8-4 package
% Sat Aug  2 17:11:21 2025
  \hline
10th & 10th & 10.5 & 11.8 & 14.1 \\ 
  10th & 50th & 11.2 & 12.3 & 14.3 \\ 
  10th & 90th & 11.8 & 12.7 & 14.5 \\ 
   \hline
50th & 10th & 9.4 & 10.9 & 13.7 \\ 
  50th & 50th & 9.9 & 11.4 & 13.8 \\ 
  50th & 90th & 10.6 & 11.9 & 14.0 \\ 
   \hline
90th & 10th & 8.1 & 9.8 & 12.5 \\ 
  90th & 50th & 8.6 & 10.1 & 12.6 \\ 
  90th & 90th & 9.2 & 10.5 & 12.7 \\ 
\end{tabular}