### 2025-12-18 output/tables/lead_by_players.tex
Updates from Overleaf

\begin{tabular}{cc|rrr}
\multicolumn{2}{c|}{Skill Percentile} & \multicolumn{3}{c}{Disengagements}\\
Battery & Runner & 0 & 1 & 2 \\
% latex table generated in R 4.5.0 by xtable 1.8-4 package
% Thu Dec 18 00:04:13 2025
  \hline
10th & 10th & 10.4 & 11.7 & 14.1 \\ 
  10th & 50th & 11.0 & 12.1 & 14.2 \\ 
  10th & 90th & 11.9 & 12.8 & 14.6 \\ 
   \hline
50th & 10th & 9.4 & 11.1 & 13.8 \\ 
  50th & 50th & 10.0 & 11.4 & 13.9 \\ 
  50th & 90th & 11.0 & 12.2 & 14.2 \\ 
   \hline
90th & 10th & 8.2 & 9.9 & 12.6 \\ 
  90th & 50th & 8.7 & 10.2 & 12.6 \\ 
  90th & 90th & 9.7 & 10.9 & 12.9 \\ 
\end{tabular}